\documentclass{beamer}

\input{settings.tex}

\newcommand{\dq} {\dot{\mathbf{q}}}


\title{Trajectory Optimization}
\subtitle{Fundamentals of Robotics, Lecture 11}
\author{by Sergei Savin}
\centering
\date{\mydate}



\begin{document}
\maketitle





\begin{frame}{State-space}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		We can define state-space coordinates $\bo{x}$ for a mechanical system $\bo{H} \ddot{\bo{q}} + \bo{C} \dot{\bo{q}} + \bo{g} = \tau$ as follows:
		
		\begin{equation}
			\bo{x} = \begin{bmatrix}
				\bo{q} \\
				\dot{\bo{q}}
			\end{bmatrix}
		\end{equation}
	
	Defining $\bo{u} = \tau$,  $\bo{S}_q = [\bo{I} \ \ \bo{0}]$ and $\bo{S}_v = [\bo{0} \ \ \bo{I}]$ we get $\bo{q} = \bo{S}_q \bo{x}$ and $\dot{\bo{q}} = \bo{S}_v \bo{x}$. The state-space dynamics becomes:
	
	
		\begin{equation}
				\dot{\bo{x}} = 
				\bo{f}(\bo{x}, \bo{u})
				=
		   \begin{bmatrix}
				\bo{S}_v \bo{x} \\
				\bo{H}^{-1} (\bo{u} - \bo{C} \bo{S}_v \bo{x} - \bo{g})
			\end{bmatrix}
		\end{equation}
	
	\end{flushleft}
\end{frame}





\begin{frame}{Trajectory planning problem}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		We know how to control a nonlinear system $\dot{\bo{x}} = 
		\bo{f}(\bo{x}, \bo{u})$ obtained from the manipulator equations, as long as we want point-to-point or trajectory following control, and the trajectory is provided. But how do we find a feasible trajectory?
		
		\bigskip
		
		Here are aspects of a typical trajectory problems:
		
		\begin{enumerate}
			\item Drive the system to the desired state $\bo{x}^*(t_f)$ by the time $t_f$.
			
			\item Respect torque limits: $|| u_i || \leq u_{max}$.
			
			\item Respect kinematic constraints: $\dot{ q }_{min} \leq \dot{ q }_i \leq \dot{ q }_{max}$.
			
			\item Respect joint limits: $q_{i, min} \leq q_i \leq q_{i, max}$.
			
		\end{enumerate}
		
	\end{flushleft}
\end{frame}





\begin{frame}{Trajectory planning as optimization}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		We can formulate the trajectory planning as an \emph{optimal control problem} (OCP):
		
		\begin{equation}
			\begin{aligned}
				& \underset{\bo{x}(t), \bo{u}(t)}{\text{minimize}}
				& & \int\limits_{t_0}^{t_f} l \left( \bo{x}(t), \bo{u}(t) \right) dt + l_f(\bo{x}(t_f)), \\
				& \text{subject to:}
				& & \dot{\bo{x}} = 
				\bo{f}(\bo{x}, \bo{u}) \\
				& & & \bo{x}_{min} \leq \bo{x}(t) \leq \bo{x}_{max} \\
				& & & \bo{u}_{min} \leq \bo{u}(t) \leq \bo{u}_{max}
			\end{aligned}
		\end{equation}
	
	This is an optimization problem with continuous variables and there are no solvers that can solve it (in a general case). So, our method is to replace the continuous time variables with a finite number of parameters. This is called \emph{transcription}.
		
	\end{flushleft}
\end{frame}



\begin{frame}{Types of transcription}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		There are a number of popular ways to transcribe the trajectory. They are often divided into \emph{collocation} and \emph{shooting} methods.
		
		\bigskip
		
		Shooting methods transcribe $\bo{u}(t)$ via finite number of parameters (spline coefficients or values of $\bo{u}$ at particular time intervals), and then compute $\bo{x}(t)$ via integration.
		
		\bigskip
		
		Collocation methods transcribe both $\bo{u}(t)$ and $\bo{x}(t)$ via finite number of parameters (spline coefficients for $\bo{u}(t)$ and $\bo{x}(t)$), and use dynamics equations as constraints.		
		
	\end{flushleft}
 \end{frame}




\begin{frame}{Direct collocation}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		Below is an example of a simple direct collocation, where the $\bo{x}(t)$ and $\bo{u}(t)$ are discretized at time nodes $t_1, \ ..., \ t_n$ as $(\bo{x}_1, \bo{u}_1), \ ..., \ (\bo{x}_n, \bo{u}_n)$.
		%
		\begin{equation*}
			\begin{aligned}
				& \underset{\substack{\bo{x}_1, \ ..., \ \bo{x}_n, \\ \bo{u}_1, \ ..., \ \bo{u}_{n-1}}}{\text{minimize}}
				& & \sum\limits_{i=1}^{n-1} \left( \bo{x}_i\T \bo{Q}_i \bo{x}_i + \bo{u}_i\T \bo{R}_i \bo{u}_i \right) + (\bo{x}_n - \bo{x}_d)\T \bo{Q}_n (\bo{x}_n - \bo{x}_d), \\
				& \text{subject to:}
				& & \frac{\bo{x}_{i+1} - \bo{x}_i}{\Delta t_i} = 
				\bo{f}(\bo{x}_i, \bo{u}_i) \\
				& & & \bo{x}_{min} \leq \bo{x}_i \leq \bo{x}_{max} \\
				& & & \bo{u}_{min} \leq \bo{u}_i \leq \bo{u}_{max}
			\end{aligned}
		\end{equation*}
		
		Approximating derivative as a difference does not strike us as highly accurate in this case. We can do a little better with the following constraint:
		%
		\begin{equation}
		\bo{x}_{i+1} = \frac{\Delta t_i}{2} (\bo{f}(\bo{x}_i, \bo{u}_i) + \bo{f}(\bo{x}_{i+1}, \bo{u}_{i+1}))
		\end{equation} 
		
	\end{flushleft}
\end{frame}




\begin{frame}{Direct collocation - linear discrete systems}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		In case of a linear discrete system $\bo{x}_{i+1} = \bo{A}_i \bo{x}_i + \bo{B}_i \bo{u}_i$, direct collocation becomes the obvious choice, equivalent to MPC. With quadratic cost it becomes a finite-time version of LQR:
		%
		\begin{equation*}
			\begin{aligned}
				& \underset{\substack{\bo{x}_1, \ ..., \ \bo{x}_n, \\ \bo{u}_1, \ ..., \ \bo{u}_{n-1}}}{\text{minimize}}
				& & \sum\limits_{i=1}^{n-1} \left( \bo{x}_i\T \bo{Q}_i \bo{x}_i + \bo{u}_i\T \bo{R}_i \bo{u}_i \right) + (\bo{x}_n - \bo{x}_d)\T \bo{Q}_n (\bo{x}_n - \bo{x}_d), \\
				& \text{subject to:}
				& & \bo{x}_{i+1} = \bo{A}_i \bo{x}_i + \bo{B}_i \bo{u}_i \\
				& & & \bo{x}_{min} \leq \bo{x}_i \leq \bo{x}_{max} \\
				& & & \bo{u}_{min} \leq \bo{u}_i \leq \bo{u}_{max}
			\end{aligned}
		\end{equation*}
		
		Note, that if there were no inequality constraints, this problem would be solved analytically.
		
	\end{flushleft}
\end{frame}




\begin{frame}{Single shooting}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		Below is an example of a simple single shooting, where the $\bo{u}(t)$ is discretized at time nodes $t_1, \ ..., \ t_n$ as $\bo{u}_1, \ ..., \  \bo{u}_n$.
		
		\begin{equation}
			\begin{aligned}
				& \underset{\substack{\bo{u}_1, \ ..., \ \bo{u}_n}}{\text{minimize}}
				& & \sum\limits_{i=1}^{n-1} ||\bo{u}_i|| + (\bo{x}(t_f) - \bo{x}_d)\T \bo{Q}_n (\bo{x}(t_f) - \bo{x}_d), \\
				& \text{subject to:}
				& & \bo{x} =  \int\limits_{t_0}^{t_f} \bo{f}(\bo{x}, \bo{u}) dt\\
				& & & \bo{u}_{min} \leq \bo{u}_i \leq \bo{u}_{max}
			\end{aligned}
		\end{equation}
		
		Note that here it is possible to use sophisticated integration schemes, but also it is hard to impose state constraints (joint limits, joint velocity limits, obstacle avoidance, etc.).
		
		
	\end{flushleft}
\end{frame}




\begin{frame}{Read more}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		
\textcolor{blue}{\href{http://www.matthewpeterkelly.com/research/MatthewKelly_IntroTrajectoryOptimization_SIAM_Review_2017.pdf}{Matthew Kelly Intro Trajectory Optimization.}}


\end{flushleft}
\end{frame}


\begin{frame}{Thank you!}
\centerline{Lecture slides are available via Moodle.}
\bigskip
\centerline{You can help improve these slides at:}
\centerline{\mygit}
\bigskip
\centerline{Check Moodle for additional links, videos, textbook suggestions.}
\bigskip

\centerline{\textcolor{black}{\qrcode[height=1.6in]{https://github.com/SergeiSa/Fundamentals-of-robotics-2022}}}

\end{frame}

\end{document}
