\documentclass{beamer}

\input{settings.tex}

\newcommand{\dq} {\dot{\mathbf{q}}}


\title{Nonlinear Control, PD, Gravity compensation, Lyapunov}
\subtitle{Fundamentals of Robotics, Lecture 9}
\author{by Sergei Savin}
\centering
\date{\mydate}



\begin{document}
\maketitle





\begin{frame}{Lyapunov method: stability criteria}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		\begin{block}{Asymptotic stability criteria}
			Autonomous dynamic system $\dot{\bo{x}} = \bo{f}(\bo{x})$ is assymptotically stable, if there exists a scalar function $V = V(\bo{x}) > 0$, whose time derivative is negative $\dot V(\bo{x}) < 0$, except $V(\bo{0}) = 0$, $\dot V(\bo{0}) = 0$.
		\end{block}
		
		\begin{block}{Marginal stability criteria}
			$\dot{\bo{x}} = \bo{f}(\bo{x})$ is stable in the sense of Lyapunov, $\exists V(\bo{x}) > 0$, $\dot V(\bo{x}) \leq 0$.
		\end{block}
		
		\begin{definition}
			Function $V(\bo{x}) > 0$ in this case is called \emph{Lyapunov function}.
		\end{definition}
		
		\bigskip
		
		This is not the only type of stability as you remember, you are invited to study criteria for other stability types on your own.
		
	\end{flushleft}
\end{frame}


\begin{frame}{Basic second-order system}
	\begin{flushleft}
		
		Let us consider the following basic system:
		
		\begin{equation}
			\ddot{\bo{q}} + \bo{K}_d \dot{\bo{q}} + \bo{K}_p\bo{q} = 0
		\end{equation}
		%
		where $\bo{K}_d > 0$ and $\bo{K}_p > 0$ are semidefinite matrices. Can we analyze its stability?
		
		\bigskip
		
		We propose Lyapunov function
		 
		 \begin{equation}
		 	V = \frac{1}{2} \dot{\bo{q}}\T  \dot{\bo{q}} + 
		 	       \frac{1}{2} \bo{q}\T \bo{K}_p\bo{q}
		 \end{equation}
		
	\end{flushleft}
\end{frame}



\begin{frame}{Basic second-order system}
	\begin{flushleft}
		
		Derivative of the Lyapunov function is:
		
		\begin{equation}
			\dot V = \dot{\bo{q}}\T  \ddot{\bo{q}} + 
			\bo{q}\T \bo{K}_p \dot{\bo{q}}
		\end{equation}
		
		But we know that $\ddot{\bo{q}} = -\bo{K}_d \dot{\bo{q}} - \bo{K}_p\bo{q}$, so:
		%
		\begin{align}
	\dot V = -\dot{\bo{q}}\T  (\bo{K}_d \dot{\bo{q}} + \bo{K}_p\bo{q}) + 
	\bo{q}\T \bo{K}_p \dot{\bo{q}} 
	\\
	\dot V = -\dot{\bo{q}}\T\bo{K}_d \dot{\bo{q}} 
		\end{align}		
	
	This means that the system is at least marginally stable: $\dot V \leq 0$, namely $\dot V = 0$ for $\dot{\bo{q}} = 0$ and $\forall \bo{q}$.
		
		\bigskip
		
		Now, we want to prove that the only trajectory in the fixed point ($\dot{\bo{q}} = 0$, $\ddot{\bo{q}} = 0$) is $\bo{q} = 0$.
		
	\end{flushleft}
\end{frame}




\begin{frame}{Basic second-order system}
	\begin{flushleft}
		
		Consider the fixed point $\dot{\bo{q}} = 0$, $\ddot{\bo{q}} = 0$:
		
		\begin{equation}
			\ddot{\bo{q}} + \bo{K}_d \dot{\bo{q}} + \bo{K}_p\bo{q} = 0 
			\ \ \ \rightarrow \ \ \ 
			\bo{K}_p\bo{q} = 0
		\end{equation}
		
		Since $\bo{K}_p > 0$, it is full rank and has a trivial null space, so $\bo{q} = 0 $ is the only trajectory in the fixed point space. So, the system is asymptotically stable by LaSalle's invariance principle.
		
	\end{flushleft}
\end{frame}





\begin{frame}{Gravity compensation}
	\begin{flushleft}
		
		Now, we can consider manipulator equations:
		
		\begin{equation}
			\bo{H} \ddot{\bo{q}} + \bo{C} \dot{\bo{q}} + \bo{g} = \tau
		\end{equation}
	
		We propose a Lyapunov function:
		 
\begin{equation}
	V = \frac{1}{2} \dot{\bo{q}}\T  \bo{H} \dot{\bo{q}} + 
	\frac{1}{2} \bo{q}\T \bo{K}_p\bo{q}
\end{equation}		
		
	\begin{align}
	\dot V = \dot{\bo{q}}\T  \bo{H}  \ddot{\bo{q}} + 
	\frac{1}{2} \dot{\bo{q}}\T  \dot{\bo{H}} \dot{\bo{q}} +
	\dot{\bo{q}}\T \bo{K}_p \bo{q} 
	\\
	\dot V = \dot{\bo{q}}\T  (\tau - \bo{C} \dot{\bo{q}} + \bo{g} ) + 
	\frac{1}{2} \dot{\bo{q}}\T  \dot{\bo{H}} \dot{\bo{q}} +
	\dot{\bo{q}}\T \bo{K}_p \bo{q} 
	\\
	\dot V = \frac{1}{2} \dot{\bo{q}}\T  ( \dot{\bo{H}} - 2 \bo{C} ) \dot{\bo{q}} +
	\dot{\bo{q}}\T  (\tau - \bo{g} ) + 
	\dot{\bo{q}}\T \bo{K}_p \bo{q} 
	\end{align}		
		
	\end{flushleft}
\end{frame}



\begin{frame}{Gravity compensation}
	\begin{flushleft}
		
		Looking at the expression: 
		
		$$\dot V = \frac{1}{2} \dot{\bo{q}}\T  ( \dot{\bo{H}} - 2 \bo{C} ) \dot{\bo{q}} +
		\dot{\bo{q}}\T  (\tau - \bo{g} ) + 
		\dot{\bo{q}}\T \bo{K}_p \bo{q} $$ 
		%
		we can observe that $ \dot{\bo{H}} - 2 \bo{C} $ is skew-symmetric, and therefore $\dot{\bo{q}}\T  ( \dot{\bo{H}} - 2 \bo{C} ) \dot{\bo{q}} = 0$
		
		We can propose the following $\tau$:
		
		\begin{equation}
			\tau = \bo{g}  - \bo{K}_d \dot{\bo{q}}  - \bo{K}_p \bo{q}
		\end{equation}		
	
		Then the Lyapunov function derivative takes form:
		
	\begin{align}
	\dot V = 
	-\dot{\bo{q}}\T  \bo{K}_d \dot{\bo{q}}
	-\dot{\bo{q}}\T  \bo{K}_p \bo{q}
	 + 
	\dot{\bo{q}}\T \bo{K}_p \bo{q} 
	\\
	\dot V = 
	-\dot{\bo{q}}\T  \bo{K}_d \dot{\bo{q}} \leq 0
	\end{align}				
					
		
	\end{flushleft}
\end{frame}




\begin{frame}{Gravity compensation}
	\begin{flushleft}
		
		So, $\dot V = 
		-\dot{\bo{q}}\T  \bo{K}_d \dot{\bo{q}} \leq 0$. More precisely $\dot V = 0$ for $\dot{\bo{q}} = 0$ and $\forall \bo{q}$, giving us marginal stability.
		
		\bigskip
		
		Let us study the fixed points $\dot{\bo{q}} = 0$ and $\ddot{\bo{q}} = 0$ for the manipulator eq.:
		
		\begin{align}
			\bo{g} = \bo{g}  - \bo{K}_p \bo{q} \\
			\bo{K}_p \bo{q} = 0 \\
			\bo{q} = 0
		\end{align}				
	
		We get that the only trajectory in the fixed space is $\bo{q} = 0$. So, the system is asymptotically stable by LaSalle's invariance principle.	
		
	\end{flushleft}
\end{frame}




\begin{frame}{Gravity compensation}
	\begin{flushleft}
		
		The control that we proposed:
		
		\begin{equation}
			\tau = \bo{g}  - \bo{K}_d \dot{\bo{q}}  - \bo{K}_p \bo{q}
		\end{equation}		
	%
	is gravity compensation and PD control. We have proven that it is  stable.
		
	\end{flushleft}
\end{frame}


\begin{frame}{Proportional control}
	\begin{flushleft}
		
		If the system has linear dissipation force (viscous force) $\bo{F} \dot{\bo{q}}$:
		
		\begin{equation}
			\bo{H} \ddot{\bo{q}} + \bo{C} \dot{\bo{q}} + \bo{F} \dot{\bo{q}} + \bo{g} = \tau
		\end{equation}
		
		...we can use proportional control without derivative component:
		
		\begin{equation}
			\tau = \bo{g}  - \bo{K}_p \bo{q}
		\end{equation}		
	
	Where the same Lyapunov function will obtain derivative:
	
	\begin{align}
		\dot V = 
		-\dot{\bo{q}}\T  \bo{F}_d \dot{\bo{q}} \leq 0
	\end{align}	
	
		With the same conclusions as before. An example of system for which we can use it is industrial robot arms with gear box reducers.
		
	\end{flushleft}
\end{frame}





\begin{frame}{Computed torque control}
	\begin{flushleft}
		
		Let us introduce  \emph{desired position} $\bo{q}^*$ and \emph{position error}  $\bo{e}$:
		
		\begin{equation}
			\bo{e} = \bo{q}^* - \bo{q}
		\end{equation}
	
		If we are lucky, the \emph{error dynamics} is evolving in accord with the following equation: 
		
		\begin{equation}
	\ddot{\bo{e}} + \bo{K}_d \dot{\bo{e}} + \bo{K}_p \bo{e} = 0
		\end{equation}		
	%
	where $\bo{K}_d > 0$ and $\bo{K}_p > 0$. This system is asymptotically stable, meaning that $\bo{e} \rightarrow 0$ and $\bo{q} \rightarrow \bo{q}^*$.
		
	\end{flushleft}
\end{frame}




\begin{frame}{Computed torque control}
	\begin{flushleft}
		
		Let us find what  $\ddot{\bo{e}}$ is:
		
		\begin{align}
			\ddot{\bo{e}} = \ddot{\bo{q}}^* - \ddot{\bo{q}} 
			\\
			\ddot{\bo{e}} = \ddot{\bo{q}}^* -  \bo{H}^{-1}(\tau - \bo{C} \dot{\bo{q}} - \bo{g})
		\end{align}
	
		Then error dynamics is:
		%
		\begin{align}
\ddot{\bo{q}}^* -  \bo{H}^{-1}(\tau - \bo{C} \dot{\bo{q}} - \bo{g})
 + \bo{K}_d \dot{\bo{e}} + \bo{K}_p \bo{e} = 0
 \\
\bo{H}\ddot{\bo{q}}^* - (\tau - \bo{C} \dot{\bo{q}} - \bo{g})
+ \bo{H}(\bo{K}_d \dot{\bo{e}} + \bo{K}_p \bo{e}) = 0 
\\
\tau = \bo{H}\ddot{\bo{q}}^* + \bo{C} \dot{\bo{q}} + \bo{g} +
\bo{H}(\bo{K}_d \dot{\bo{e}} + \bo{K}_p \bo{e})
		\end{align}				
	
	So, we found a control law that makes error dynamics stable, meaning $\bo{q} \rightarrow \bo{q}^*$.
		
	\end{flushleft}
\end{frame}




\begin{frame}{Read more}
	% \framesubtitle{Parameter estimation}
	\begin{flushleft}
		
		You can read more at \emph{ Siciliano, B., Sciavicco, L., Villani, L. and Oriolo, G., 2009. Robotics. Advanced textbooks in control and signal processing}, Chapter 8.5
		

\end{flushleft}
\end{frame}


\begin{frame}{Thank you!}
\centerline{Lecture slides are available via Moodle.}
\bigskip
\centerline{You can help improve these slides at:}
\centerline{\mygit}
\bigskip
\centerline{Check Moodle for additional links, videos, textbook suggestions.}
\bigskip

\centerline{\textcolor{black}{\qrcode[height=1.6in]{https://github.com/SergeiSa/Fundamentals-of-robotics-2022}}}

\end{frame}

\end{document}
